---
title: "Logboek"
author: "Kasper Notebomer"
date: "9/14/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library("ggplot2")
library("kableExtra")
library("dplyr")
library(tidyr)
library(purrr)
library(ggcorrplot)
library(reshape2)
```

<!-- * Copyright (c) 2018 Kasper Notebomer. -->
<!-- * Licensed under GPLv3. See gpl.md -->



## Introduction


## Data exploration

To start off we'll read the given data set using the read.csv function. After that the data structure and the five number summary is viewed to look for any problems with the data that was read in.

```{r}
data <- read.csv("data/dataFrame_all_sims.csv", sep = ",", header=TRUE)
str(data)
```
```{r}
sum <- summary(data[c("temperature","sterol.conc","satur.index","PC.conc","ethanol.conc","APL", "thickness", "bending", "tilt", "zorder","compress")])
sum 
sum <- sub(".*:", "", sum)
rownames(sum) <- c("Minimum", "Q1", "Median", "Mean", "Q3", "Maximum", "Number of NA's")

kable(sum) %>%
  kable_styling(latex_options = "scale_down")
```
Looking at the structure there doesn't seem to be immediate problem with the data, like the data being read as a wrong type of data. There do seem to be some rows that are missing values like the APL, thickness, beding and tilt values. These rows will most likely need to be removed.

```{r}
chart <- ggplot(data, aes(temperature, APL, colour = factor(satur.index), shape = factor(ethanol.conc) )) + 
         geom_point(alpha=1/10) + geom_jitter()
chart
```
```{r}
chart2 <- ggplot(data, aes(APL, bending, colour = factor(satur.index), shape = factor(sterol.type) )) + 
        geom_jitter(alpha=1)
chart2
```
```{r}
p <- ggplot(data, aes(x=bending, colour = factor(sterol.type))) + 
  geom_density()
p
```
```{r}
p <- ggplot(data, aes(x=APL, colour = factor(tails))) + 
  geom_density()
p
```
```{r}
cor2 <- data[,9:14]
cor2 <- na.omit(cor2)
cormat <- round(cor(cor2),2)
head(cormat)

melted_cormat <- melt(cormat)
head(melted_cormat)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```



Here a codebook is made of all of the abreviations in the dataframe and their meaning.
```{r}
colnames(data)
codebook <- data.frame("Abreviation" = colnames(data), "Type/Class" = sapply(data, class), 
                       "Desciption" = c("Temperature (Kelvin)", "Sterol type", 
                                        "Sterol concentration (%)", "Other (phospho)lipids in membrane (headgroup)",
                                        "Aliphatic tails", "Saturation index (double bonds per tail)", 
                                        "Phosphatidyl choline concentration (% of non-sterol lipids)", 
                                        "Ethanol concentration (% of solvent)", "Area per lipid (nm^2)", 
                                        "Thickness (nm)", "Bending rigidity (kB T)", "Tilt angle (degrees)", 
                                        "Z-order", "Compressibility (cN / m)"))
rownames(codebook) <- NULL
codebook
sapply(data, typeof)

write.csv(codebook, "data/codebook.csv", row.names = FALSE)
```


Onderzoeksvraag?
Is it possible to use machine learning to predict the bending regidity of a membrane given the composition of this membrane.
Is it possible to use machine learning to predict the bending regidity of a membrane given the sterol type and sterol consentration of the membrane.

different dataset:
Is is possible to use machine learning to identify a bird based on their song?
